---
title: "S&P 500 Index Time Series EDA"
author: "C. Roberts/R. J. Serrano"
format: 
    html:
        code-tools: true
        self-contained: true
editor: visual
toc: true
execute: 
  cache: true
---

Source: [Exploratory Data Analysis of Time Series Data](https://medium.com/@mrconnor/exploratory-data-analysis-of-time-series-data-987f8a0c2a0a)

```{r}
#| echo: false
suppressPackageStartupMessages({
    library(tidyverse)
    library(timetk)
    library(patchwork)
})

theme_set(theme_minimal())
```

Read the S&P 500 Index Adjusted Closing, Daily return (%) and volatility dataset.

```{r}
spx <- read_rds('../data/spx_return_vol_tbl.rds')

spx
```

## Plot time series

::: panel-tabset
### S&P 500 Index Adjusted Closing

```{r}
spx %>% 
     select(date, adjusted) %>% 
     plot_time_series(date, adjusted, 
                      .title = 'S&P 500 Index Daily Adjusted Closing Price (Jan 2013 - Jan 2023 (partial)')
```

### S&P 500 Index Daily Returns (%)

```{r}
spx %>% 
     plot_time_series(date, return, 
                      .title = 'S&P 500 Index Daily Adjusted Closing Return Percentage (Jan 2013 - Jan 2023 (partial))')
```

### S&P 500 Index Daily Volatility (%)

```{r}
spx %>% 
     plot_time_series(date, volatility, 
                      .title = 'S&P 500 Index Daily Volatility Percentage (Jan 2013 - Jan 2023 (partial))')
```
:::

## ACF/PACF Diagnostics

::: panel-tabset
### S&P 500 Index Adjusted Closing

```{r}
#| fig-height: 8
spx %>% 
     plot_acf_diagnostics(date, adjusted)
```

### S&P 500 Index Daily Returns (%)

```{r}
#| fig-height: 8
spx %>% 
     plot_acf_diagnostics(date, return)
```

### S&P 500 Index Daily Volatility (%)

```{r}
#| fig-height: 8
spx %>% 
     plot_acf_diagnostics(date, volatility)
```

:::

## Seasonal Diagnostics

::: panel-tabset
### S&P 500 Index Adjusted Closing

```{r}
#| fig-height: 10
spx %>% 
     plot_seasonal_diagnostics(
          .date_var = date, 
          .value = adjusted
     )
```

### S&P 500 Index Daily Returns (%)

```{r}
#| fig-height: 10
spx %>% 
     plot_seasonal_diagnostics(
          .date_var = date, 
          .value = return
     )
```

### S&P 500 Index Daily Volatility (%)

```{r}
#| fig-height: 10
spx %>% 
     plot_seasonal_diagnostics(
          .date_var = date, 
          .value = volatility
     )
```

:::

## Anomaly Diagnostics

::: panel-tabset
### S&P 500 Index Adjusted Closing

```{r}
spx %>% 
     plot_anomaly_diagnostics(
          .date_var = date, 
          .value = adjusted, 
          .alpha = 0.05, 
          .max_anomalies = 0.03
     )
```

### S&P 500 Index Daily Returns (%)

```{r}
spx %>% 
     plot_anomaly_diagnostics(
          .date_var = date, 
          .value = return, 
          .alpha = 0.05, 
          .max_anomalies = 0.03
     )
```

### S&P 500 Index Daily Volatility (%)

```{r}
spx %>% 
     plot_anomaly_diagnostics(
          .date_var = date, 
          .value = volatility, 
          .alpha = 0.05, 
          .max_anomalies = 0.03
     )
```

:::

## Seasonal Decomposition

::: panel-tabset
### S&P 500 Index Adjusted Closing

```{r}
#| fig-height: 10
spx %>% 
     plot_stl_diagnostics(
          .date_var = date, 
          .value = adjusted
     )
```

### S&P 500 Index Daily Returns (%)

```{r}
#| fig-height: 10
spx %>% 
     plot_stl_diagnostics(
          .date_var = date, 
          .value = return
     )
```

### S&P 500 Index Daily Volatility (%)

```{r}
#| fig-height: 10
spx %>% 
     plot_stl_diagnostics(
          .date_var = date, 
          .value = volatility
     )
```

:::

## Heteroskedasticity (variance not uniform across the time series) test

Using [bptest](https://en.wikipedia.org/wiki/Breusch–Pagan_test) from the `lmtest` package

Hypothesis test:

- Null hypothesis (H~0~): Time series variance is uniform

- Alternate hypothesis (H~a~): Time series variance is not uniform

::: panel-tabset
### S&P 500 Index Adjusted Closing

```{r}
#| echo: false
library(lmtest)
```

```{r}
lm_model_adj <- lm(adjusted ~ as.numeric(date), data = spx)

bptest(lm_model_adj, data = spx)
```

Since the p-value < 0.05, we can reject the null hypothesis in favor of the alternate hypothesis, i.e., their is significant evidence that the time series variance is not uniform (may require transformation).

### S&P 500 Index Daily Returns (%)

```{r}
lm_model_ret <- lm(return ~ as.numeric(date), data = spx)

bptest(lm_model_ret, data = spx)
```

Since the p-value < 0.05, we can reject the null hypothesis in favor of the alternate hypothesis, i.e., their is significant evidence that the time series variance is not uniform (may require transformation).

### S&P 500 Index Daily Volatility (%)

```{r}
lm_model_vol <- lm(volatility ~ as.numeric(date), data = spx)

bptest(lm_model_ret, data = spx)
```

Since the p-value < 0.05, we can reject the null hypothesis in favor of the alternate hypothesis, i.e., their is significant evidence that the time series variance is not uniform (may require transformation).

:::

## Stationarity Test

What is the definition of a stationary time series?

According to the textbook [Chapter 8.1 - Stationarity and differencing](https://otexts.com/fpp2/stationarity.html), *"a stationary time series is one whose properties do not depend on the time at which the series is observed. Thus, time series with trends, or with seasonality, are not stationary — the trend and seasonality will affect the value of the time series at different times."*

Is there a test to detect time series stationarity?

Yes. The traditional test is the ADF (Augmented Dick Fuller) test.

Hypothesis test:

- Null hypothesis (H~0~): Time series is non-stationary

- Alternate hypothesis (H~a~): Time series is stationary

::: panel-tabset
### S&P 500 Index Adjusted Closing

```{r}
#| echo: false
library(tseries)
```

```{r}
adf.test(spx$adjusted)
```

Since the p-value > 0.05, we cannot reject the null hypothesis. Therefore, the time series is non-stationary.

### S&P 500 Index Daily Returns (%)

```{r}
adf.test(spx$return)
```

Since the p-value < 0.05, we can reject the null hypothesis in favor of the alternate hypothesis, i.e., the time series is stationary.

### S&P 500 Index Daily Volatility (%)

```{r}
adf.test(spx$volatility)
```

Since the p-value < 0.05, we can reject the null hypothesis in favor of the alternate hypothesis, i.e., the time series is stationary.

:::